\textbf{Mammalian protein-coding genes exhibit widespread beneficial mutations that are not adaptive}

### 2024-12-27 manuscript/previous-versions/supp-mat-PNAS-revision-2.tex
Moved tex files into a single folder

    \newpage
    \section{Gene ontology enrichment}

    For each gene, we computed the proportion of $\SphyBen$ sites.
    For each ontology, the proportion of sites under adaptation is compared between the set of genes sharing this given ontology and the rest of the genes with Mann-Whitney U test.
    $p_{\mathrm{v}}^{\mathrm{adj}}$ are corrected for multiple comparison (Holm–Bonferroni correction).
    $^*$ for $p_{\mathrm{v}}^{\mathrm{adj}}$ corrected for multiple comparison (Holm–Bonferroni correction) lower than the risk $\alpha=0.05$.
        \captionof{table}{Ontology enrichment}
        \begin{longtable*}{|l|r|r|r|r|r|r|}
            Gene ontology & p(go) & r(go/no go) & Mann-Whitney U & $p_{\mathrm{v}}$ & $p_{\mathrm{v}}^{\mathrm{adj}}$ \\
            \multicolumn{6}{r}{Continued on next page} \\
            proteinaceous extracellul... & $ 0.069$ & $ 1.679$ & $2.4\times 10^{5}$ & $0.00042$ & $ 0.118~~$ \\
            extracellular region & $ 0.178$ & $ 1.518$ & $ 8\times 10^{5}$ & $ 0.002$ & $ 0.599~~$ \\
            extracellular matrix orga... & $ 0.039$ & $ 2.102$ & $1.3\times 10^{5}$ & $ 0.003$ & $ 0.746~~$ \\
            receptor-mediated endocyt... & $ 0.027$ & $ 1.253$ & $8.1\times 10^{4}$ & $ 0.006$ & $ 1.000~~$ \\
            regulation of transcripti... & $ 0.069$ & $ 0.482$ & $7.1\times 10^{5}$ & $ 0.008$ & $ 1.000~~$ \\
            cytosol & $ 0.212$ & $ 0.714$ & $1.4\times 10^{6}$ & $ 0.013$ & $ 1.000~~$ \\
            nucleoplasm & $ 0.124$ & $ 0.505$ & $ 1\times 10^{6}$ & $ 0.014$ & $ 1.000~~$ \\
            transcription from RNA po... & $ 0.008$ & $ 0.106$ & $2.3\times 10^{5}$ & $ 0.015$ & $ 1.000~~$ \\
            microtubule-based movemen... & $ 0.023$ & $ 1.054$ & $7.2\times 10^{4}$ & $ 0.016$ & $ 1.000~~$ \\
            nucleus & $ 0.266$ & $ 0.713$ & $1.5\times 10^{6}$ & $ 0.020$ & $ 1.000~~$ \\
            ATPase activity & $ 0.031$ & $ 0.924$ & $9.3\times 10^{4}$ & $ 0.025$ & $ 1.000~~$ \\
            mitochondrion & $ 0.058$ & $ 0.517$ & $5.5\times 10^{5}$ & $ 0.029$ & $ 1.000~~$ \\
            transcription & $ 0.073$ & $ 0.544$ & $6.8\times 10^{5}$ & $ 0.030$ & $ 1.000~~$ \\
            cytoplasm & $ 0.317$ & $ 0.708$ & $1.6\times 10^{6}$ & $ 0.032$ & $ 1.000~~$ \\
            Golgi membrane & $ 0.012$ & $ 0.102$ & $2.4\times 10^{5}$ & $ 0.033$ & $ 1.000~~$ \\
            RNA polymerase II transcr... & $ 0.027$ & $ 0.460$ & $3.4\times 10^{5}$ & $ 0.046$ & $ 1.000~~$ \\
            cell adhesion & $ 0.062$ & $ 1.220$ & $2.9\times 10^{5}$ & $ 0.054$ & $ 1.000~~$ \\
            actin filament binding & $ 0.019$ & $ 0.897$ & $6.7\times 10^{4}$ & $ 0.060$ & $ 1.000~~$ \\
            dendrite & $ 0.008$ & $ 0.112$ & $1.7\times 10^{5}$ & $ 0.067$ & $ 1.000~~$ \\
            regulation of signal tran... & $ 0.019$ & $ 1.402$ & $ 7\times 10^{4}$ & $ 0.069$ & $ 1.000~~$ \\
            sequence-specific DNA bin... & $ 0.015$ & $ 0.286$ & $2.1\times 10^{5}$ & $ 0.072$ & $ 1.000~~$ \\
            spermatogenesis & $ 0.039$ & $ 1.633$ & $1.6\times 10^{5}$ & $ 0.081$ & $ 1.000~~$ \\
            cell surface receptor sig... & $ 0.031$ & $ 2.126$ & $1.3\times 10^{5}$ & $ 0.086$ & $ 1.000~~$ \\
            regulation of receptor ac... & $ 0.027$ & $ 3.275$ & $1.2\times 10^{5}$ & $ 0.088$ & $ 1.000~~$ \\
            Golgi apparatus & $ 0.046$ & $ 0.477$ & $4.8\times 10^{5}$ & $ 0.094$ & $ 1.000~~$ \\
            receptor complex & $ 0.023$ & $ 1.407$ & $9.4\times 10^{4}$ & $ 0.108$ & $ 1.000~~$ \\
            positive regulation of tr... & $ 0.039$ & $ 0.435$ & $4.1\times 10^{5}$ & $ 0.111$ & $ 1.000~~$ \\
            viral process & $ 0.008$ & $ 0.844$ & $1.5\times 10^{5}$ & $ 0.114$ & $ 1.000~~$ \\
            mRNA processing & $ 0.004$ & $ 0.061$ & $1.1\times 10^{5}$ & $ 0.118$ & $ 1.000~~$ \\
            endocytosis & $ 0.023$ & $ 1.602$ & $9.8\times 10^{4}$ & $ 0.123$ & $ 1.000~~$ \\
            transferase activity & $ 0.004$ & $ 0.089$ & $ 1\times 10^{5}$ & $ 0.125$ & $ 1.000~~$ \\
            negative regulation of ce... & $ 0.015$ & $ 3.268$ & $6.2\times 10^{4}$ & $ 0.125$ & $ 1.000~~$ \\
            positive regulation of GT... & $ 0.035$ & $ 2.351$ & $1.7\times 10^{5}$ & $ 0.127$ & $ 1.000~~$ \\
            oxidation-reduction proce... & $ 0.019$ & $ 0.754$ & $2.6\times 10^{5}$ & $ 0.128$ & $ 1.000~~$ \\
            post-translational protei... & $ 0.008$ & $ 0.299$ & $1.4\times 10^{5}$ & $ 0.130$ & $ 1.000~~$ \\
            mitochondrial inner membr... & $ 0.008$ & $ 0.661$ & $1.5\times 10^{5}$ & $ 0.130$ & $ 1.000~~$ \\
            protein C-terminus bindin... & $ 0.004$ & $ 0.021$ & $ 1\times 10^{5}$ & $ 0.134$ & $ 1.000~~$ \\
            transcription factor bind... & $ 0.004$ & $ 0.159$ & $ 1\times 10^{5}$ & $ 0.135$ & $ 1.000~~$ \\
            cytoskeleton organization & $ 0.019$ & $ 1.607$ & $7.9\times 10^{4}$ & $ 0.137$ & $ 1.000~~$ \\
            cell migration & $ 0.004$ & $ 0.110$ & $9.9\times 10^{4}$ & $ 0.143$ & $ 1.000~~$ \\
            extracellular space & $ 0.100$ & $ 1.295$ & $5.4\times 10^{5}$ & $ 0.144$ & $ 1.000~~$ \\
            extracellular exosome & $ 0.104$ & $ 0.694$ & $8.4\times 10^{5}$ & $ 0.145$ & $ 1.000~~$ \\
            apoptotic process & $ 0.015$ & $ 0.660$ & $2.2\times 10^{5}$ & $ 0.145$ & $ 1.000~~$ \\
            cell junction & $ 0.031$ & $ 0.314$ & $3.4\times 10^{5}$ & $ 0.152$ & $ 1.000~~$ \\
            DNA binding & $ 0.077$ & $ 0.705$ & $6.5\times 10^{5}$ & $ 0.155$ & $ 1.000~~$ \\
            postsynaptic density & $ 0.004$ & $ 0.223$ & $9.7\times 10^{4}$ & $ 0.157$ & $ 1.000~~$ \\
            hydrolase activity & $ 0.081$ & $ 0.526$ & $6.7\times 10^{5}$ & $ 0.157$ & $ 1.000~~$ \\
            transcription regulatory ... & $ 0.004$ & $ 0.228$ & $9.5\times 10^{4}$ & $ 0.163$ & $ 1.000~~$ \\
            DNA binding transcription... & $ 0.027$ & $ 0.436$ & $3.1\times 10^{5}$ & $ 0.164$ & $ 1.000~~$ \\
            signal transducer activit... & $ 0.027$ & $ 0.371$ & $ 3\times 10^{5}$ & $ 0.167$ & $ 1.000~~$ \\
            aging & $ 0.015$ & $ 2.107$ & $6.3\times 10^{4}$ & $ 0.169$ & $ 1.000~~$ \\
            transcription coactivator... & $ 0.004$ & $ 0.687$ & $9.6\times 10^{4}$ & $ 0.172$ & $ 1.000~~$ \\
            synapse & $ 0.019$ & $ 0.255$ & $2.3\times 10^{5}$ & $ 0.172$ & $ 1.000~~$ \\
            positive regulation of ER... & $ 0.019$ & $ 2.201$ & $8.4\times 10^{4}$ & $ 0.172$ & $ 1.000~~$ \\
            sensory perception of sou... & $ 0.015$ & $ 1.737$ & $6.3\times 10^{4}$ & $ 0.182$ & $ 1.000~~$ \\
            metalloendopeptidase acti... & $ 0.019$ & $ 0.600$ & $ 8\times 10^{4}$ & $ 0.183$ & $ 1.000~~$ \\
            response to drug & $ 0.004$ & $ 0.082$ & $8.8\times 10^{4}$ & $ 0.183$ & $ 1.000~~$ \\
            skeletal system developme... & $ 0.015$ & $ 1.590$ & $6.5\times 10^{4}$ & $ 0.189$ & $ 1.000~~$ \\
            mitochondrial matrix & $ 0.008$ & $ 0.421$ & $1.3\times 10^{5}$ & $ 0.193$ & $ 1.000~~$ \\
            immune system process & $ 0.012$ & $ 0.299$ & $1.6\times 10^{5}$ & $ 0.194$ & $ 1.000~~$ \\
            cell-matrix adhesion & $ 0.015$ & $ 0.999$ & $6.3\times 10^{4}$ & $ 0.198$ & $ 1.000~~$ \\
            receptor binding & $ 0.031$ & $ 2.044$ & $1.5\times 10^{5}$ & $ 0.201$ & $ 1.000~~$ \\
            ion transmembrane transpo... & $ 0.008$ & $ 0.363$ & $1.3\times 10^{5}$ & $ 0.204$ & $ 1.000~~$ \\
            nervous system developmen... & $ 0.015$ & $ 0.401$ & $ 2\times 10^{5}$ & $ 0.205$ & $ 1.000~~$ \\
            oxidoreductase activity & $ 0.019$ & $ 0.843$ & $2.3\times 10^{5}$ & $ 0.210$ & $ 1.000~~$ \\
            innate immune response & $ 0.012$ & $ 0.243$ & $1.6\times 10^{5}$ & $ 0.218$ & $ 1.000~~$ \\
            RNA polymerase II proxima... & $ 0.012$ & $ 0.595$ & $1.6\times 10^{5}$ & $ 0.219$ & $ 1.000~~$ \\
            positive regulation of ce... & $ 0.004$ & $ 0.117$ & $ 8\times 10^{4}$ & $ 0.230$ & $ 1.000~~$ \\
            protein binding & $ 0.587$ & $ 0.735$ & $1.7\times 10^{6}$ & $ 0.244$ & $ 1.000~~$ \\
            binding & $ 0.046$ & $ 0.648$ & $ 2\times 10^{5}$ & $ 0.248$ & $ 1.000~~$ \\
            DNA replication & $ 0.019$ & $ 1.388$ & $8.9\times 10^{4}$ & $ 0.248$ & $ 1.000~~$ \\
            RNA splicing & $ 0.008$ & $ 0.095$ & $7.6\times 10^{4}$ & $ 0.250$ & $ 1.000~~$ \\
            transcription factor comp... & $ 0.004$ & $ 0.280$ & $7.8\times 10^{4}$ & $ 0.253$ & $ 1.000~~$ \\
            transmembrane transport & $ 0.054$ & $ 1.414$ & $2.8\times 10^{5}$ & $ 0.255$ & $ 1.000~~$ \\
            neuronal cell body & $ 0.012$ & $ 0.386$ & $1.5\times 10^{5}$ & $ 0.255$ & $ 1.000~~$ \\
            regulation of cell shape & $ 0.015$ & $ 1.945$ & $ 7\times 10^{4}$ & $ 0.256$ & $ 1.000~~$ \\
            external side of plasma m... & $ 0.004$ & $ 0.042$ & $7.5\times 10^{4}$ & $ 0.259$ & $ 1.000~~$ \\
            endomembrane system & $ 0.015$ & $ 3.003$ & $7.4\times 10^{4}$ & $ 0.264$ & $ 1.000~~$ \\
            protein folding & $ 0.015$ & $ 1.691$ & $7.2\times 10^{4}$ & $ 0.267$ & $ 1.000~~$ \\
            protein polyubiquitinatio... & $ 0.004$ & $ 0.431$ & $7.7\times 10^{4}$ & $ 0.271$ & $ 1.000~~$ \\
            inflammatory response & $ 0.027$ & $ 1.745$ & $1.4\times 10^{5}$ & $ 0.273$ & $ 1.000~~$ \\
            protein heterodimerizatio... & $ 0.035$ & $ 1.731$ & $1.9\times 10^{5}$ & $ 0.283$ & $ 1.000~~$ \\
            membrane raft & $ 0.008$ & $ 0.207$ & $1.1\times 10^{5}$ & $ 0.285$ & $ 1.000~~$ \\
            transcriptional activator... & $ 0.008$ & $ 0.242$ & $1.1\times 10^{5}$ & $ 0.290$ & $ 1.000~~$ \\
            perinuclear region of cyt... & $ 0.027$ & $ 0.611$ & $2.5\times 10^{5}$ & $ 0.291$ & $ 1.000~~$ \\
            regulation of cell prolif... & $ 0.019$ & $ 1.644$ & $9.4\times 10^{4}$ & $ 0.291$ & $ 1.000~~$ \\
            sodium ion transport & $ 0.004$ & $ 0.185$ & $7.1\times 10^{4}$ & $ 0.294$ & $ 1.000~~$ \\
            negative regulation of ce... & $ 0.015$ & $ 0.395$ & $1.8\times 10^{5}$ & $ 0.304$ & $ 1.000~~$ \\
            macromolecular complex bi... & $ 0.004$ & $ 0.190$ & $ 7\times 10^{4}$ & $ 0.307$ & $ 1.000~~$ \\
            positive regulation of tr... & $ 0.023$ & $ 0.782$ & $2.4\times 10^{5}$ & $ 0.315$ & $ 1.000~~$ \\
            cilium & $ 0.023$ & $ 0.956$ & $1.2\times 10^{5}$ & $ 0.315$ & $ 1.000~~$ \\
            signal transduction & $ 0.089$ & $ 0.724$ & $ 7\times 10^{5}$ & $ 0.317$ & $ 1.000~~$ \\
            lipid metabolic process & $ 0.046$ & $ 1.542$ & $2.4\times 10^{5}$ & $ 0.321$ & $ 1.000~~$ \\
            cell-cell signaling & $ 0.004$ & $ 0.475$ & $ 7\times 10^{4}$ & $ 0.325$ & $ 1.000~~$ \\
            protein homooligomerizati... & $ 0.004$ & $ 0.475$ & $ 7\times 10^{4}$ & $ 0.325$ & $ 1.000~~$ \\
            integrin binding & $ 0.004$ & $ 0.143$ & $6.6\times 10^{4}$ & $ 0.331$ & $ 1.000~~$ \\
            actin binding & $ 0.027$ & $ 0.693$ & $1.4\times 10^{5}$ & $ 0.344$ & $ 1.000~~$ \\
            symporter activity & $ 0.004$ & $ 0.334$ & $6.6\times 10^{4}$ & $ 0.348$ & $ 1.000~~$ \\
            DNA repair & $ 0.015$ & $ 0.277$ & $1.7\times 10^{5}$ & $ 0.351$ & $ 1.000~~$ \\
            cell differentiation & $ 0.066$ & $ 0.834$ & $3.5\times 10^{5}$ & $ 0.358$ & $ 1.000~~$ \\

    \newpage
    \section{Clinically related terms for mutations}

    \subsection{Terms associated with deleterious mutations $\SphyDel$}
    \begin{center}
        \captionof{table}{Terms associated with deleterious mutations $\SphyDel$.}
        \begin{tabular}{|l|r|r|r|r|r|}
            \toprule
            SNP clinical ontology & $n_{\mathrm{Observed}}$ & $n_{\mathrm{Expected}}$ & Odds ratio & $p_{\mathrm{v}}$ & $p_{\mathrm{v-adjusted}}$ \\
            \midrule
            Benign                & 2969                    & $4043.0$                & $ 0.734$   & $ 1.000$             & $ 1.000~~$                    \\
            Likely benign         & 2994                    & $3399.8$                & $ 0.881$   & $ 0.999$             & $ 1.000~~$                    \\
            Risk factor           & 102                     & $ 118.2$                & $ 0.863$   & $ 0.798$             & $ 1.000~~$                    \\
            Likely pathogenic     & 221                     & $  68.5$                  & $ 3.226$   & $1.7\times 10^{-8}$  & $\bm{6.7\times 10^{-8}{^*}}$  \\
            Pathogenic            & 560                     & $ 193.6$                & $ 2.893$   & $4.2\times 10^{-17}$ & $\bm{2.1\times 10^{-16}{^*}}$ \\
            \bottomrule
        \end{tabular}
    \end{center}
    In humans, non-synonymous SNPs in the test group ($\SphyDel$) are contrasted to SNPs in the control group ($\SphyNeu$).
    For each clinical term, a 2x2 contingency tables is built by counting the number of SNPs based on their selection coefficient and their clinical terms (whether they have this specific term or not).
    Fisher's exact tests are then performed for these 2x2 contingency tables.
    $^*$ for $p_{\mathrm{v}}^{\mathrm{adj}}$ corrected for multiple comparison (Holm–Bonferroni correction) lower than the risk $\alpha=0.05$.
    SNPs predicted with $\SphyDel$ are statistically associated to clinical terms such as \textit{Likely Pathogenic} and \textit{Pathogenic}.

>>>>

    \section{Clinically related terms for mutations}
    \subsection{Terms associated with deleterious mutations $\SphyDel$}
    \begin{center}
        \captionof{table}{Terms associated with deleterious mutations $\SphyDel$.}
        \begin{tabular}{|l|r|r|r|r|r|}
            \toprule
            SNP clinical ontology & $n_{\mathrm{Observed}}$ & $n_{\mathrm{Expected}}$ & Odds ratio & $p_{\mathrm{v}}$ & $p_{\mathrm{v-adjusted}}$ \\
            \midrule
            Benign                & 2969                    & $4043.0$                & $ 0.734$   & $ 1.000$             & $ 1.000~~$                    \\
            Likely benign         & 2994                    & $3399.8$                & $ 0.881$   & $ 0.999$             & $ 1.000~~$                    \\
            Risk factor           & 102                     & $ 118.2$                & $ 0.863$   & $ 0.798$             & $ 1.000~~$                    \\
            Likely pathogenic     & 221                     & $  68.5$                  & $ 3.226$   & $1.7\times 10^{-8}$  & $\bm{6.7\times 10^{-8}{^*}}$  \\
            Pathogenic            & 560                     & $ 193.6$                & $ 2.893$   & $4.2\times 10^{-17}$ & $\bm{2.1\times 10^{-16}{^*}}$ \\
            \bottomrule
        \end{tabular}
    \end{center}
    In humans, non-synonymous SNPs in the test group ($\SphyDel$) are contrasted to SNPs in the control group ($\SphyNeu$).
    For each clinical term, a 2x2 contingency tables is built by counting the number of SNPs based on their selection coefficient and their clinical terms (whether they have this specific term or not).
    Fisher's exact tests are then performed for these 2x2 contingency tables.
    $^*$ for $p_{\mathrm{v}}^{\mathrm{adj}}$ corrected for multiple comparison (Holm–Bonferroni correction) lower than the risk $\alpha=0.05$.
    SNPs predicted with $\SphyDel$ are statistically associated to clinical terms such as \textit{Likely Pathogenic} and \textit{Pathogenic}.
